### 2018-03-08 paper/background.tex
Work on Background

Build systems automate the execution of simple repeatable tasks for individual
users, as well as for large organisations. There are software build systems such
as \Make, \Ninja, \Shake, \Bazel, \Buck and many others, as well as various
incremental calculation engines, such as \Excel and \Calc. Existing build
systems vary in several different ways:

>>>>

Build systems automate the execution of simple repeatable tasks for individual
users, as well as for large organisations. There are software build systems,
such as \Make~\cite{feldman1979make}, \Shake~\cite{mitchell2012shake} and
\Bazel~\cite{bazel}, as well as various incremental calculation engines, such
as \Excel~\cite{advanced_excel}. In this section we use these four examples to
introduce main domain-specific notions and requirements. Other notable examples
of build systems and their relation to these four will be discussed
in~\S\ref{sec-related} and~\S\ref{sec-conclusions}.

\Make\footnote{There are numerous implementations of \Make and none comes with a
formal specification. In this paper we therefore use a simple and sensible
approximation to a real \Make that you might find on your machine.} was developed
more than 40 years ago to automatically build software libraries and executable
programs from source code. It uses \emph{makefiles} to describe build tasks,
typically referred to as \emph{build rules}, and their dependencies in a simple
text form. For example:

\begin{minted}[frame=single]{makefile}
util.o: util.h util.c
    gcc -c util.c

main.o: util.h main.c
    gcc -c main.c

main.exe: util.o main.o
    gcc util.o main.o -o main.exe
\end{minted}

\noindent
The above makefile lists three tasks: (i) compile a utility library comprising
files \cmd{util.h} and \cmd{util.c} into \cmd{main.o} by executing the command
\cmd{gcc -c util.c}, (ii) compile the main source file \cmd{main.c} into
\cmd{main.o}, and (iii) link object files \cmd{util.o} and \cmd{main.o} into the
executable \cmd{main.exe}. The makefile contains the complete information about
the task \emph{dependency graph}, which is shown in Fig.~\ref{fig-make}(a).

\begin{figure}[h]
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example.pdf}}
\caption{A dependency graph}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-full.pdf}}
\caption{A full rebuild}
\end{subfigure}
\begin{subfigure}[b]{0.32\linewidth}
\centerline{\includegraphics[scale=0.28]{fig/make-example-partial.pdf}}
\caption{A partial rebuild}
\end{subfigure}
\caption{A dependency graph and two build scenarios. Input files are shown as
rectangles, intermediate and output files are shown as rounded rectangles. Dirty
inputs and files that are rebuilt are highlighted.
\label{fig-make}}
\end{figure}

If the user modifies the sources of the utility library and runs \Make, it will
perform a full rebuild, because all three tasks transitively depend on the
library, as illustrated in Fig.~\ref{fig-make}(b). On the other hand, if the
user modifies \cmd{main.c} then a partial rebuild is sufficient: indeed, the
file \cmd{util.o} does not need to be rebuilt, since its inputs have not
changed, see Fig.~\ref{fig-make}(c). Files that have changed since the previous
build are called \emph{dirty}.

The requirement to execute tasks \emph{at most once} and only if they
\emph{transitively depend on dirty inputs} is essential for build systems,
their raison d'\^etre. We will call build systems that satisfy this requirement
\emph{minimal}.

To achieve minimality \Make relies on two key ideas. First, it uses \emph{file
modification time} to detect the files that are dirty: a file is marked dirty if
it was modified after the previous build. Second, it constructs a complete task
dependency graph from the information contained in the makefile and executes
tasks in the \emph{topological order}.

\Excel

\Shake

\begin{figure}[h]
\centerline{\includegraphics[scale=0.28]{fig/make-example-cutoff.pdf}}
\caption{An early cut-off example.\label{fig-cutoff}}
\end{figure}

\Bazel
